---
title: "Things you need to know about Christmas gift-giving"
date: "2020-12-24"
categories: ["R"]
tags: ["R Blogdown", "R Markdown", "ggplot2"]
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, out.width = "100%")
library(ggplot2)
library(stringr)
library(magrittr)
library(reshape2)
library(sysfonts)
font_add_google("Source Sans Pro")
```

It's the most wonderful time of the year and the happiest season of all, [Andy Williams](https://www.youtube.com/watch?v=gFtb3EtjEic) sings so. But for people currently dating someone, buying a Christmas present may seem not only like a huge responsibility, but also like an incredibly tough choice. 

Once the pressure on holiday traditions hits, everyone's anxiety goes higher with the list of gifts to be bought. Some people spend a lot of time searching and researching for the perfect gift, in the hope that it will bring that sparkle and joy to our beloved ones. And a lot of things differ when we're talking about a present for your partner, starting with how long you have been dating or going out. Let me give you some bad news first.

#### Unless you've been going out for a few weeks already, it's unlikely that you'll get a present from your "Santa".

[YouGov](https://yougov.co.uk/topics/arts/survey-results/daily/2019/11/26/882a1/2) surveyed 1329 britains to find out what people think about when you should get your date a Christmas present. It seems like men are twice as likely than women to buy a gift after only 1-2 dates, but when it comes to longer relationships, women are almost twice more convinced that they should keep alive the Christmas spirit and offer a gift to their partner. 

Actually, women are more reluctant to buying gifts early in the relationship, but after going out for a few weeks, they're much more open to Christmas gift-giving. The remaining attitudes are quite similar across genders, but it's clear that unless you've been going out for a few weeks already, it's unlikely that you'll get a present from your "Santa". 

```{r when-to-buy-xmas-presents, tidy = FALSE, echo = FALSE, message = FALSE, warning = FALSE}
# When to buy Christmas presents data
xmas_labels <- c(#"Don't know",
                 "You have been going out for longer than a few months", 
                 "You have been going out for a few months", 
                 "You have been going out for a few weeks", 
                 "You have been on 3-4 dates but are not yet going out",
                 "You have been on 1-2 dates", 
                 "You are due to go on your 1st date around Christmas time"
                 ) %>% factor(., levels = .)
xmas_male.values <- c(#.3, 
  .05, .12, .27, .09, .11, .06)
xmas_female.values <- c(#.25, 
  .09, .2, .3, .07, .05, .04)
xmas.data <- data.frame(Reason = xmas_labels, 
                        Male = xmas_male.values, 
                        Female = xmas_female.values)
xmas.df <- reshape2::melt(xmas.data, id.vars = "Reason") %>% 
  magrittr::set_names(c("Reason", "Gender", "Value"))

# When to buy a present - ggplot
ggplot(data = xmas.df, aes(x = Reason, y = Value, fill = Gender)) +
  geom_bar(stat = "identity", position = position_dodge(-.9)) +
  scale_x_discrete(labels = stringr::str_wrap(xmas_labels, width = 25), 
                   name = element_blank()) + 
  scale_y_continuous(name = element_blank()) + 
  geom_text(aes(x = Reason, y = Value, fill = Gender, 
                label = paste0(Value*100, "%")), 
            colour = "white", size = 3.3, fontface = "bold", 
            position = position_dodge(-.9), 
            hjust = +1.3, vjust = +.45) + 
  coord_flip() +
  scale_fill_manual(values = c("#28543C", # Christmas green
                               "#901A1A"  # Christmas Red
                               ), aesthetics = "fill") + 
  theme_minimal() +   
  theme(text = element_text(family = "Source Sans Pro"),
        plot.title = element_text(hjust = 0.5, face = "bold", size = 15),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 10),
        panel.grid = element_blank(), 
        axis.line.y.left = element_line(colour = "grey", size = 0.2),
        legend.title = element_blank(),
        legend.position = c(.9, .9),
        strip.text = element_blank()) + 
  ggtitle(label = stringr::str_wrap("You should you get a Christmas present if...",
                                    width = 50)) +
  guides(guide_legend(reverse = TRUE)) + 
  facet_wrap(~ Gender)
```

Interestingly, the percentages drop for people who have been going out for longer than a few months. As a matter of fact, men are slightly more likely to say they should get a Christmas present at the first date rather than for a committed partner. It feels like men consider the Chrismas gift-giving more of an early relationship thing that fades away with time, whereas women associate offering gifts with a long-term relationship.

#### Safest choice for everyone: gift cards. 

Finding the right gift may give a lot of headaches and stress to some people. If you still find yourself in the position of looking for a gift idea, here's what people would like to receive for Chrismas, according to [Statista](https://www.statista.com/statistics/643724/christmas-gifts-desired-by-consumers-by-gender/). Offering a gift card seems like the best suitable option for everyone, especially under the current pandemic restrictions. Don't worry, it's never late for getting a gift coupon, even if it's already Christmas Eve. 

Next options differ slightly by gender. Women prefer to receive money, clothing, cosmetics or jewelry - or just returnable products? They're also more willing to say they want books, so if you've identified a lady as a heavy reader, you will never find a better gift for her than a book! 

```{r what-to-buy-for-xmas, tidy = FALSE, echo = FALSE, message = FALSE, warning = FALSE}
# Gift preferences for Christmas - data
xmas_gift_ideas <- c("Games, toys",
                     #"DIY and home improvement products",
                     #"None",
                     #"Furniture and decoration",
                     #"Household appliances",
                     "Consumer electronics, media, video and audio",
                     "Mobile phones, tablets, accessories",
                     "Computers and accessories",
                     "Food, liquor",
                     "Jewelery and watches",
                     "Cosmetics, fragrances, heath & beauty aids",
                     "Books",
                     "Clothing, textiles, shoes",
                     "Money",   
                     "Gift cards or gift certificates"
                     ) %>% factor(., levels = .)

xmas_gift_male.values <- c(.16, #.13, .17, .15, .13,
  .30, .20, .31, .24, .17, .17, .23, .33, .23, .36)
xmas_gift_female.values <- c(.08, #.11, .12, .13, .13, 
  .17, .18, .18, .18, .22, .26, .27, .36, .36, .51)
xmas_gift.data <- data.frame(Preference = xmas_gift_ideas, 
                             Male = xmas_gift_male.values, 
                             Female = xmas_gift_female.values) 
xmas_gift.df <- reshape2::melt(xmas_gift.data, id.vars = c("Preference")) %>% 
  magrittr::set_names(c("Preference", "Gender", "Value"))

# Gift preferences for Christmas - ggplot
ggplot(data = xmas_gift.df, aes(x = Preference, y = Value, fill = Gender)) +
  geom_bar(stat = "identity", position = position_dodge(-.9), guide = FALSE) +
  scale_x_discrete(labels = stringr::str_wrap(xmas_gift_ideas, width = 25), 
                   name = element_blank()) + 
  scale_y_continuous(name = element_blank()) +
  geom_text(aes(x = Preference, y = Value, fill = Gender, 
                label = paste0(Value*100, "%")), 
            colour = "white", size = 3.3, fontface = "bold", 
            position = position_dodge(-.9), 
            hjust = +1.3, vjust = +.45) + 
  coord_flip() +
  scale_fill_manual(values = c("#28543C", # Christmas green
                               "#901A1A"  # Christmas Red
                               ), aesthetics = "fill") + 
  theme_minimal() +   
  theme(text = element_text(family = "Source Sans Pro"),
        plot.title = element_text(hjust = 0.5, face = "bold", size = 15),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 10),
        panel.grid = element_blank(), 
        axis.line.y.left = element_line(colour = "grey", size = 0.2),
        strip.text = element_blank(),
        legend.title = element_blank(),
        legend.position = c(.9, .08)) + 
  ggtitle(label = stringr::str_wrap("Which types of gifts would you like to receive for Christmas?",
                                    width = 60)) +
  facet_wrap(~ Gender)
```

At the opposite pole, men prefer to receive clothing, computers and accessories or media eletronics - once a techie, always a techie! Otherwise, choosing a book or some liquor for the spirits lovers could be a better idea.

Some men are twice more likely to wish for games and toys, although very few people actually fall into this category (less than 2 in 10 men). An avid gamer is definitely into more games! 

Now that you know all of that, it should be much easier for you to choose an appropriate gift for your partner. In the Christmas Eve, everyone deserves that joy given by receiving a present, no matter if they were naughty or nice. And don't forget: spending quality time with people is always more valuable than any gift we can offer.

Sources: [YouGov, 2017](https://yougov.co.uk/topics/arts/survey-results/daily/2019/11/26/882a1/2); [Statista, 2017](https://www.statista.com/statistics/643724/christmas-gifts-desired-by-consumers-by-gender/).